---
title: "DATA 606 Data Project Proposal"
author: "Richard Zheng"
output: pdf_document
---

### Packages

```{r}
library(tidyverse)
```

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data
url = 'https://raw.githubusercontent.com/schoolkidrich/housing_price/main/train.csv'

data = read.csv(url, stringsAsFactors = TRUE)
```

### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

#### What factors are relevant in determining house price? 

### Cases 

**What are the cases, and how many are there?**

```{r}
# each row represents one house 
num_cases = length(data[,1])
print(paste("There are",num_cases,"cases"))
```


### Data collection 

**Describe the method of data collection.**

#### Data is obtained from kaggle as part of one of their competitions


### Type of study 

**What type of study is this (observational/experiment)?**

#### This is an observational study

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

#### 'https://www.kaggle.com/c/house-prices-advanced-regression-techniques/overview'


### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

#### Dependant variable are the housing prices. It is quantitative 

### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**

```{r}
#There are many variables in this dataset (both quantitative and qualitative ones). I will try to use as many as possible 
variables = length(names(data))-2

print(paste("There are",variables,"variables"))

```



### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
summary(data)
```

# Lot Size vs Price

```{r}
data %>% 
  ggplot(aes(x = SalePrice,y = LotArea))+geom_point(size = 2)+ylim(0,50000)
```

# Paved vs Gravel

```{r}
data %>%
  ggplot(aes(x = Street, y = SalePrice))+geom_boxplot()
```



```{r}
drop = c('Id', 'Alley', 'Heating', 'Condition1', 'Condition2', 'BsmtHalfBath','X3SsnPorch', 'PoolQC','MiscFeature','Fence', 'LotArea', 'Utilities', 'RoofMatl','GarageTypeDetchd', 'TotalBsmtSF', 'GrLivArea', 'GarageType')
df = data[,!(names(data) %in% drop)]

model1 = lm(SalePrice~., df)
summary(model1)
```

